Glyph-maps for visualizing climate data and models

### 2011-08-26 paper.tex
tweaking intro

Climate data is composed of measurements on variables such as temperature, precipitation, and winds that have been collected from a spatial and temporal context. The classic display for spatial data is to use color or grey scale overlaid on geographic location, typically including a map. When measurements are made a different time periods also a common way to display the data is in small multiples, such as in Figure \ref{fig:facetted-map}. In this plot, a separate map is drawn for each month (columns) and year (rows), over a 6 year period of remotely sensed temperature data above Central America. Color is used to display the de-seasonalized temperature measurements, with red being high values and blue being low values. The data  comes from \cite{NASA-Data-Expo}. The most noticeable feature in this display is the strong red in the equatorial region beginning in mid-1997 and tapering out during 1998. This is the pattern of an El Ni\~no event, a major temperature anomaly. With some work other smaller more localized patterns can be perceived, for example, cooler over the land mass in the early year, but warmer in the latter years.

>>>>

Climate data is composed of measurements on variables such as temperature, precipitation, and winds that have been collected from a spatial and temporal context. The classic display for spatial data is to use color or grey scale overlaid on geographic location, typically including a map. When measurements are made a different time periods also a common way to display the data is in small multiples, such as in Figure \ref{fig:facetted-map}. In this plot, a separate map is drawn for each month (columns) and year (rows), over a 6 year period of remotely sensed temperature data above Central America. Color is used to display the de-seasonalized temperature measurements, with red being high values and blue being low values. The data  comes from \cite{NASA-Data-Expo}. The most noticeable feature in this display is the strong red in the equatorial region beginning in mid-1997 and tapering out during 1998. This is the pattern of an El Ni\~no event, a major temperature anomaly. With some work other smaller more localized patterns can be perceived, for example, cooler over the land mass in the early year, but warmer in the latter years. What the reader is required to do is play the game "Spot the difference" from one small image to another. It's a very hard task for readers to do. Large structure such as the El Ni\~no event are clear but readers will not be able to effectively glue the images together sufficiently to read off the trend over the years or notice small local deviations between plots.
